---
title: "Water, and its Factors in different Countries"
author: "- Ryan Louie (rlouie21@uw.edu)
- Christopher Chen (cmchen22@uw.edu)
- Sean Lee (seanl6@uw.edu)
- Mynah Shetty (mshett@uw.edu)"
date: "11/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(data.table)
library(knitr)

```

```{r load_data}
  #Cholera data
cholera <- read.csv("../data/data_cholera_infection_rate.csv")

#Handwashing data
handwash <- read.csv("../data/data_handwash.csv")

#Population data
population <- read.csv("../data/data_country_population.csv")

#Diarrhea data
diarrhea <- read.csv("../data/data_diarrhea_death_rate.csv")

#Poverty data
poverty <- read.csv("../data/data_poverty.csv")

handwash <- read.csv("../data/data_handwash.csv")
```
## Dynamic Paragraph

```{r, echo = FALSE, message=FALSE}
  #Convert Years to rows
population_filtered <- select(population, -c(Country.Code, Indicator.Name, Indicator.Code)) %>%
                      gather("Year", "Population", -Country.Name)%>%
                      rename(Country =Country.Name)
#Rename columns for handwash
handwash_filtered <- slice(handwash,3:n())
names(handwash_filtered) <-  paste0(names(handwash_filtered), ",") %>%
                               str_replace("\\.1,", ", Urban") %>%
                          str_replace("\\.2,", ", Rural") %>%
                          str_replace("\\,$", ", Both Rural and Urban")
#Remove Years to actual years
population_filtered$Year <- strtoi(str_replace(population_filtered$Year, "X", ""))
#Convert handwash years to rows
handwash_filtered <- rename(handwash_filtered, Countries = "X, Both Rural and Urban") %>%
                      gather("Year_and_type", "Percent", -Countries) %>%
                    separate("Year_and_type", into = c("Year","Urban_Rural_or_Both"), sep = ", ")%>%
                    rename(Country = Countries)
#Convert Percentages to integers
#String to integer function is from:
#https://www.tutorialkart.com/r-tutorial/r-convert-string-to-integer/#:~:text=To%20convert%20String%20to%20Integer%20in%20R%20programming%2C%20call%20strtoi,respect%20to%20the%20specified%20base.
handwash_filtered$Percent <- strtoi(handwash_filtered$Percent)
handwash_filtered <- filter(handwash_filtered, !is.na(handwash_filtered$Percent))
#Convert Year to actual years
handwash_filtered$Year <- strtoi(str_replace(handwash_filtered$Year, "X", ""))
#Join handwash_filtered and cholera and filter for max date
handwash_and_cholera <- inner_join(handwash_filtered, cholera) %>% filter(Year == max(Year))
#Value 1
maximum_proportion_handwash <- handwash_and_cholera %>%filter(Percent == max(Percent)) %>%
  pull(Percent)
maximum_proportion_handwash <- maximum_proportion_handwash[1]
#Value 2
max_prop_how_many <-  handwash_and_cholera %>% filter(Percent == max(Percent)) %>%
  pull(Country)
max_prop_how_many <- length(max_prop_how_many)
#Value 3
avg_prop_poverty <- round((mean(na.omit(poverty$percPoverty))), digits = 2)
#Value 4
median_cholera_cases  <- round(median(handwash_and_cholera$Number.of.reported.cases.of.cholera, na.rm =TRUE), digits = 2)
#Value 5
minimum_proportion_handwash <-  handwash_and_cholera %>% filter(Percent == min(Percent)) %>%
  pull(Percent)
minimum_proportion_handwash <- minimum_proportion_handwash[1]
#Value 6
min_prop_how_many <-  handwash_and_cholera %>% filter(Percent == min(Percent)) %>%
  pull(Country)
min_prop_how_many <- length(min_prop_how_many)
#Summary list function
summary_list <- list()
summary_list$maximum_proportion_handwash <- maximum_proportion_handwash
summary_list$max_prop_how_many <- max_prop_how_many
summary_list$avg_prop_poverty <- avg_prop_poverty
summary_list$median_cholera_cases <- median_cholera_cases
summary_list$minimum_proportion_handwash <-minimum_proportion_handwash
summary_list$min_prop_how_many <- min_prop_how_many
summary_list$avg_cholera_cases <- round(mean(handwash_and_cholera$Number.of.reported.cases.of.cholera, na.rm =TRUE), digits = 2)
```
## Table
```{r, echo = FALSE}
highest_poverty <- poverty %>%
  group_by(country)%>%
  filter(percPoverty >= 50.0)
table <- as.data.table(highest_poverty)
kable(table)
 
```

I chose this table because it represents the countries that have a majority poverty population. I included this table because we can use the information off this table to determine the link between poverty and disease. The information presented on this table include the name of the country, the percentage of poverty in each country, the year the data was collected, and the percentage of population that learn less than 10 dollars a day, and less than 5.50 dollars a day specifically. This table allows us to see the most impovirished countries in the world, which we can use to establish conenctions between widespread economic struggle, and increase of disease rates. 


## Chart 1
```{r, echo = FALSE, message = FALSE}
  #Convert Years to rows
population_filtered <- select(population, -c(Country.Code, Indicator.Name, Indicator.Code)) %>%
                      gather("Year", "Population", -Country.Name)%>%
                      rename(Country =Country.Name)
#Rename columns for handwash
handwash_filtered <- slice(handwash,3:n())
names(handwash_filtered) <-  paste0(names(handwash_filtered), ",") %>%
                               str_replace("\\.1,", ", Urban") %>%
                          str_replace("\\.2,", ", Rural") %>%
                          str_replace("\\,$", ", Both Rural and Urban")

#Remove Years to actual years
population_filtered$Year <- strtoi(str_replace(population_filtered$Year, "X", ""))

#Convert handwash years to rows
handwash_filtered <- rename(handwash_filtered, Countries = "X, Both Rural and Urban") %>%
                      gather("Year_and_type", "Percent", -Countries) %>%
                    separate("Year_and_type", into = c("Year","Urban_Rural_or_Both"), sep = ", ")%>%
                    rename(Country = Countries)

#Convert Percentages to integers
#String to integer function is from:
#https://www.tutorialkart.com/r-tutorial/r-convert-string-to-integer/#:~:text=To%20convert%20String%20to%20Integer%20in%20R%20programming%2C%20call%20strtoi,respect%20to%20the%20specified%20base.
handwash_filtered$Percent <- strtoi(handwash_filtered$Percent)
handwash_filtered <- filter(handwash_filtered, !is.na(handwash_filtered$Percent))

#Convert Year to actual years
handwash_filtered$Year <- strtoi(str_replace(handwash_filtered$Year, "X", ""))

#Join handwash_filtered and cholera
handwash_and_cholera <- inner_join(handwash_filtered, cholera)

#Join handwash_and_cholera with population_filtered, and
#create Proportion for Cholera Cases divided by Proportion of Sanitation
final_dataframe <- inner_join(handwash_and_cholera, population_filtered) %>%
                    mutate(cases_per_population = Number.of.reported.cases.of.cholera / Population * 100) %>%
                    filter(Year == 2016 | Year == 2015 | Year == 2014 | Year ==2013 | Year == 2012 | Year == 2011| Year == 2010 )

#Make our ggplot2 chart
chart_1 <- ggplot(data = final_dataframe, aes(x = Percent, y = cases_per_population)) +
                    geom_point(alpha = .4, color = "seagreen") +
                    geom_smooth()+facet_wrap(~Urban_Rural_or_Both) +
                    xlab("Percentage of People with Access to Clean Services") +
                    ylab("Reported Cases of Cholera per Population of Country") +
                  # I got the scales::comma lable from:
                  #https://stackoverflow.com/questions/52602503/display-an-axis-value-in-millions-in-ggplot
                  #Square root scale is from:
                  #https://ggplot2-book.org/scales-guides.html
                    scale_y_sqrt(labels = scales::comma) +
                    labs(title = "Percent of Access to Clean Services vs Percent of Reported Cases of Cholera per Country Population",
                        subtitle = "Relationship between clean services and reported cases from 2010 to 2016",
                        caption = "This chart was created to access the relation
                        between the percentage of people in a country with
                        access to clean handwashing services and the reported cases of cholera per population of country.
                        Our initial belief was that there would be a negative correlation
                        between the two variables as more access to clean handwashing services
                        reduces the risk of diseases. The graph suggests a non-linear
                        relationship, but with a general negative trend.")


chart_1


```





## Chart 2
```{r, echo = FALSE, message = FALSE}
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)



handwash_filtered <- slice(handwash,3:n())
names(handwash_filtered) <-  paste0(names(handwash_filtered), ",") %>%
  str_replace("\\.1,", ", Urban") %>%
  str_replace("\\.2,", ", Rural") %>%
  str_replace("\\,$", ", Both Rural and Urban")

#Convert handwash years to rows
handwash_filtered <- rename(handwash_filtered, Countries = "X, Both Rural and Urban") %>%
  gather("Year_and_type", "Percent", -Countries) %>%
  separate("Year_and_type", into = c("Year","Urban_Rural_or_Both"), sep = ", ")%>%
  rename(Country = Countries)

#Convert Percentages to integers
#String to integer function is from:
#https://www.tutorialkart.com/r-tutorial/r-convert-string-to-integer/#:~:text=To%20convert%20String%20to%20Integer%20in%20R%20programming%2C%20call%20strtoi,respect%20to%20the%20specified%20base.
handwash_filtered$Percent <- strtoi(handwash_filtered$Percent)
handwash_filtered <- filter(handwash_filtered, !is.na(handwash_filtered$Percent))

#Convert Year to actual years
handwash_filtered$Year <- strtoi(str_replace(handwash_filtered$Year, "X", ""))

(mean(na.omit(poverty$percPoverty)))


handwash_filtered <- slice(handwash,3:n())
names(handwash_filtered) <-  paste0(names(handwash_filtered), ",") %>%
  str_replace("\\.1,", ", Urban") %>%
  str_replace("\\.2,", ", Rural") %>%
  str_replace("\\,$", ", Both Rural and Urban")

#Convert handwash years to rows
handwash_filtered <- rename(handwash_filtered, Countries = "X, Both Rural and Urban") %>%
  gather("Year_and_type", "Percent", -Countries) %>%
  separate("Year_and_type", into = c("Year","Urban_Rural_or_Both"), sep = ", ")%>%
  rename(Country = Countries)

#Convert Percentages to integers
#String to integer function is from:
#https://www.tutorialkart.com/r-tutorial/r-convert-string-to-integer/#:~:text=To%20convert%20String%20to%20Integer%20in%20R%20programming%2C%20call%20strtoi,respect%20to%20the%20specified%20base.
handwash_filtered$Percent <- strtoi(handwash_filtered$Percent)
handwash_filtered <- filter(handwash_filtered, !is.na(handwash_filtered$Percent))

#Convert Year to actual years
handwash_filtered$Year <- strtoi(str_replace(handwash_filtered$Year, "X", ""))

(mean(na.omit(poverty$percPoverty)))

poverty_df <- data.frame(poverty$percPoverty, poverty)
poverty_df <- data.frame(poverty_df$percPoverty, poverty)
poverty_df$percPoverty <- as.double(poverty_df$percPoverty)


handwash_df <- handwash_filtered %>% mutate(no = (100 - handwash_filtered$Percent)) 
View(handwash_df)

poverty.df <- poverty_df %>% mutate(
  country_type = ifelse(poverty_df$percPoverty  < 10,"rich", ifelse(poverty_df$percPoverty < 20 & poverty_df$percPoverty > 10,  
                                                                    "average", ifelse(poverty_df$percPoverty < 30 & poverty_df$percPoverty > 20, "slightly poor", ifelse(poverty_df$percPoverty < 40 & 
                                                                                                                                                                           poverty_df$percPoverty > 30, "poor", "extremely poor"))))
)

poverty.df <- poverty.df %>% 
  rename("Country" = "country")

merge <- inner_join(poverty.df, handwash_df)

View(merge)
View(poverty.df)

merge <- merge %>%group_by(country_type) %>%
  summarise(Percent = mean(Percent))
merge <- arrange(merge, Percent)
View(merge)

merge %>%filter(!is.na(country_type))

p<-ggplot(data= na.omit(merge), aes(x=country_type, y=Percent)) +
  geom_bar(stat="identity", fill = " dark green") +
  
  xlab("Country Condition based on Poverty Rates") +
  ylab("Percentage of the Country that has Access to Handwashing") +
  scale_y_sqrt(labels = scales::comma)  +
  labs(title = "Percent of Handwashing Availibility in types of Countries", 
       subtitle = "Relationship between the average handwashing percentage and country types from 2000 to 2020", 
       caption = "This depicts the average proportion of handwashing in countries based on their classification of their living quality. 
       This was calculated based on what percentage the country was in poverty." )
p


poverty_df <- data.frame(poverty$percPoverty, poverty)
poverty_df <- data.frame(poverty_df$percPoverty, poverty)
poverty_df$percPoverty <- as.double(poverty_df$percPoverty)


handwash_df <- handwash_filtered %>% mutate(no = (100 - handwash_filtered$Percent)) 
View(handwash_df)

poverty.df <- poverty_df %>% mutate(
country_type = ifelse(poverty_df$percPoverty  < 10,"rich", ifelse(poverty_df$percPoverty < 20 & poverty_df$percPoverty > 10,  
    "average", ifelse(poverty_df$percPoverty < 30 & poverty_df$percPoverty > 20, "slightly poor", ifelse(poverty_df$percPoverty < 40 & 
    poverty_df$percPoverty > 30, "poor", "extremely poor"))))
 )

poverty.df <- poverty.df %>% 
  rename("Country" = "country")

merge <- inner_join(poverty.df, handwash_df)

View(merge)
View(poverty.df)

merge <- merge %>%group_by(country_type) %>%
summarise(Percent = mean(Percent))
merge <- arrange(merge, Percent)
View(merge)

merge %>%filter(!is.na(country_type))

p<-ggplot(data= na.omit(merge), aes(x=country_type, y=Percent)) +
  geom_bar(stat="identity", fill = " dark green") +
 
  xlab("Country Condition based on Poverty Rates") +
  ylab("Percentage of the Country that has Access to Handwashing") +
   scale_y_sqrt(labels = scales::comma)  +
  labs(title = "Percent of Handwashing Availibility in types of Countries", 
       subtitle = "Relationship between the average handwashing percentage and country types from 2000 to 2020", 
       caption = "This depicts the average proportion of handwashing in countries based on their classification of their living quality. 
       This was calculated based on what percentage the country was in poverty." )
p




```


## Chart 3


The reason for working on this chart is to delve into how poverty can affect access to clean water services. This problem that stems from poverty can result in diseases such as cholera, or symptoms of diarrhea, and other factors. This may also result in health problems, and thus more money spent on those services, continuing and increasing the cycle of poverty. From the graph and data that I analyzed, the companies with the lowest poverty rates ended up having the most access to handwashing. The divide between the average of the lowest poverty rate countries and highest poverty rates was 86.17% compared to 25.21% which is a really big gap. Thus, there needs to be a lot more emphasis on countries that have high poverty rates. The data was not surprising in terms of the condition of the countries and their percentages, but the gap between each condition was alarming to see, especially at the extreme ends of each side.
